---
title: "Coursera & JHU: Reproductible Research Week 2 Peer-graded Assignment"
author: "Ana Guardia"
date: "24 de febrero de 2018"
output: html_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 0: Loading and preprocessing the data

```{r dataset: df}
df <- read.csv("activity.csv", header = TRUE, sep = ",")

head(df)
str(df)
summary(df)

```


## Part 1: What is mean total number of steps taken per day?

### 1. Calculate the total number of steps taken per day

```{r steps per day}

spd<-aggregate(steps~date,FUN=sum,data=df)
summary(spd)

```

```{r plot steps per day, echo=FALSE}

xspd<-spd$date
yspd<-spd$steps
plot(xspd,yspd,main="Total number of steps taken by day",type="n")
lines(xspd,yspd,type="l")

```

### 2. Make a histogram of the total number of steps taken each day

```{r hist steps per day, echo=FALSE}

hist(yspd, main="Histogram of steps taken by day",xlab="steps taken by day")

```

### 3. Calculate and report the mean and median of the total number of steps taken per day

```{r mean & median}

mean(yspd)
median(yspd)

```

*Note: For this part of the assignment, you can ignore the missing values in the dataset.*

## Part 2: What is the average daily activity pattern?

### Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r average steps by interval}
spi<-aggregate(steps~interval,FUN=mean,data=df)
summary(spi)
```


```{r plot steps per interval, echo=FALSE}
xspi<-spi$interval
yspi<-spi$steps
plot(xspi,yspi,main="Average number of steps by interval",type="n",xlab="interval",ylab="mean steps")
lines(xspi,yspi,type="l")
```


### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r max search}
spi[yspi==max(yspi),]
```

## Part 3: Imputing missing values

*Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.*

### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r NA count}
sum(is.na(df$steps))
```

### 2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

I will use the mean for that 5-minute interval. 

### 4. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r NA treatment}
dfNA<-df
dfNA$steps[is.na(df$steps)] <- spi$steps[match(df$interval,spi$interval)][which(is.na(df$steps))]
head(dfNA)
```

Check
````{r check NA}
sum(is.na(dfNA$steps))
```

### 5. Make a histogram of the total number of steps taken each day.
```{r hist steps per day without NAs, echo=FALSE}

yna<-dfNA$steps
hist(yna, main="Histogram of steps taken by day",xlab="steps taken by day")

```

### Calculate and report the mean and median total number of steps taken per day. 

```{r mean & median, without NAs}

mean(yna)
median(yna)

```
### Do these values differ from the estimates from the first part of the assignment? 

Yes, they do.

### What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r compare steps per day, echo=FALSE}
spd<-aggregate(steps~date,FUN=sum,data=df)
spdNA<-aggregate(steps~date,FUN=sum,data=dfNA)

xspd<-spd$date
yspd<-spd$steps
xspdNA<-spdNA$date
yspdNA<-spdNA$steps

plot(xspd,yspd,main="Total number of steps taken by day",type="n")
lines(xspd,yspd,type="l",col="blue")
lines(xspdNA,yspdNA,type="l",col="red")
```

### Part 4: Are there differences in activity patterns between weekdays and weekends?

For this part the `weekdays()` function may be of some help here. Use
the dataset with the filled-in missing values for this part.

1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r weekdays}
Sys.setlocale("LC_ALL","English")

dfNA$date <- as.Date(dfNA$date)
#create a vector of weekdays
weekdays <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
#Use `%in%` and `weekdays` to create a logical vector
#convert to `factor` and specify the `levels/labels`
dfNA$wDay <- factor((weekdays(dfNA$date) %in% weekdays), 
                    levels=c(FALSE, TRUE), labels=c('weekend', 'weekday')) 
summary(dfNA)
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r compare week and weekend days}
msteps<-aggregate(dfNA$steps, list(dfNA$interval,dfNA$wDay), mean)
colnames(msteps)[1]<-"interval"
colnames(msteps)[2]<-"wDay"
colnames(msteps)[3]<-"meansteps"
summary(msteps)
```

```{r plot to compare week and weekend days, echo=FALSE}
library(ggplot2)
qplot(interval,meansteps, data = msteps, facets=wDay~.)+ geom_line(size=2) + theme(legend.position="none")
```


## This is the End